Tree size, microhabitat, and hydraulic traits shape drought responses in a temperate broadleaf forest

### 2019-09-18 manuscript/climate_sensitivity_draft.Rmd
more edits to intro

One fundamental question regarding forest responses to drought is what drives the observed tendency for large trees to suffer more during drought? @bennett_larger_2015 showed that in forests globally, large trees suffer greater growth reductions during drought, and numerous subsequent studies have reinforced this finding (Stovall et al. 2019, **REFS**). However, this analysis quantified tree size based on DBH, which has no direct mechanistic meaning. This study proposed two major mechanisms--besides the tendency for bark beetles to preferentially attack larger trees (**REFS**)--for the observed greater drought growth reductions of large trees. First, [: (1)] inherently greater biophysical challenge of being tall]. Second, larger trees may have lower drought resistance because they are more often in the canopy, where they are exposed to higher solar radiation, greater wind speeds, and lower humidity (**REFS**). Alternatively, the generally supressed status of subcanopy trees may be insufficient to override the benefits of their buffered environment during drought. Potentially counteracting the biophysical challenges faced by large trees, their larger root systems may confer an advantage in terms of allowing greater access to water (**REFS**); however, it appears that this effect is usually insufficient to offset the costs of height and/or crown exposure. Elucidating the mechanisms that place larger trees at a relative disadvantage during drought will be critical to better understanding and modeling drought impacts on forests.

*Second, how do species' traits - alone and in interaction with tree size - influence drought response?*

@@

Moreover, there may be an interaction between traits and tree size. It is possible that the pattern observed by @bennett_larger_2015 could be caused by smaller trees being more drought resistant. Alternatively, larger trees may have more drought-resistant traits as adpatations to greater biophysical challenges.

*Third, are responses similar or variable across individual drought years?*

>>>>

**One fundamental question regarding forest responses to drought is what drives the observed tendency for large trees to suffer more during drought?** @bennett_larger_2015 showed that in forests globally, large trees suffer greater growth reductions during drought, and numerous subsequent studies have reinforced this finding (Stovall et al. 2019, **REFS**). However, this analysis quantified tree size based on DBH, which has no direct mechanistic meaning. This study proposed two major mechanisms--besides the tendency for bark beetles to preferentially attack larger trees (**REFS**)--for the observed greater drought growth reductions of large trees. First, [: (1)] inherently greater biophysical challenge of being tall]. Second, larger trees may have lower drought resistance because they are more often in the canopy, where they are exposed to higher solar radiation, greater wind speeds, and lower humidity (**REFS**). Alternatively, the generally supressed status of subcanopy trees may be insufficient to override the benefits of their buffered environment during drought. Potentially counteracting the biophysical challenges faced by large trees, their larger root systems may confer an advantage in terms of allowing greater access to water (**REFS**); however, it appears that this effect is usually insufficient to offset the costs of height and/or crown exposure. A final mechanism that could mediate tree size-related responses to drought is how hydraulic traits are distributed with respect to size ([Meakem et al. 2018](https://nph.onlinelibrary.wiley.com/doi/10.1111/nph.14633)). It is possible that the pattern observed by @bennett_larger_2015 could be caused by smaller trees being more *drought resistant*. Alternatively, larger trees may have more *drought-resistant* [**careful on wording here**] traits as adpatations to greater biophysical challenges.

**A second fundamental question regarding forest responses to drought is how species' traits - alone and in interaction with tree size - influence drought response.**

@@

(interaction between traits and tree size).

**A final fundamental question regarding forest responses to drought is whether tree size and species' traits have similar influence across droughts, or whether drought variability in factors such as severity, duration, and timing is linked to a diversity of tree repsonses. **
